
---
    title: "Practical Machine Learning Course Project - October 2016"
---
    
##Executive Summary
Prediction of the manner in which an exercise was performed using a training
data set.
Six young health participants were asked to perform one set of 10 repetitions 
of the Unilateral Dumbbell Biceps Curl in five different fashions: 
exactly according to the specification (Class A) 
throwing the elbows to the front (Class B)
lifting the dumbbell only halfway (Class C)
lowering the dumbbell only halfway (Class D) 
throwing the hips to the front (Class E)

http://groupware.les.inf.puc-rio.br/har

##Import the data
```{r}

training<-read.csv("./training.csv",na.strings=c("#DIV/0!","NA"))

testing<-read.csv("./testing.csv",na.strings=c("#DIV/0!","NA"))

str(training)
```

##Data Cleaning
Remove columns that have a large number of NAs. Removed ID columns

```{r}
training_subset<-training[,apply(training, 2, function(x) !any(is.na(x)))]
training_clean<-training_subset[,-(1:7)]


testing_subset<-testing[,apply(testing, 2, function(x) !any(is.na(x)))]
testing_clean<-testing_subset[,-(1:7)]
```

##Create partitions for Cross Validation
Used the functionality in the Caret package to split the training data set into
a training partition and a test partition to be used for cross validation of 
the model
```{r}
library(caret) #unifying machine learning package
inTrain  <- createDataPartition(
    training_clean$classe, 
    p=0.70, 
    list=FALSE)

part_train <- training_clean[inTrain, ]
part_test  <- training_clean[-inTrain, ]
```


##Random Forest Model
A Random Forest Model was chosen due to its high accuracy.
Random forests are a way of averaging multiple deep decision trees, trained 
on different parts of the same training set, with the goal of reducing the 
variance. They can be difficult to interpret but are often very accurate.

```{r}

modelRF<-train( classe~.,
    data=part_train, 
    method="rf",
    ntree=10)

modelRF

```
##Cross Validation on part_test
```{r}

predictRF<-predict(modelRF,newdata=part_test)

```
##Confusion Matrix
```{r}

confusion_matrix <- confusionMatrix(predictRF, part_test$classe)
confusion_matrix

```
###Out of Sample Error is estimated at: 1.41%


##Predictions on Test Data for Quiz
```{r}
predictQuiz<-predict(modelRF,newdata=testing_clean)

predictQuiz
```
##Conclusion

The model created was accurate enough to predict all of the 20 validation tests
correctly. Only 10 decision trees were used in this project in order to cut down
the running time of the analysis but the accuracy could be increased by 
increasing the number of trees using the ntree parameter.


